Supporting Information

### 2018-03-30 SupportingInformation.Rmd
fixed minor latex problems that prevented knitting

<!--Show the statistical results in a table-->
```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,c(3,2,4)]
colnames(s)<-c("Statistic","Lower 95\\% CI", "Mean", "Upper 95\\% CI")
rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Psq_l - \\Psq_u$","$\\Dsq_u - \\Dsq_l$")
knitr::kable(s,
             format = "pandoc",
             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
             booktabs=TRUE)
```

@@

## Aphid first flight data for 'Leaf-curling plum aphid' \label{Ex_ff}
<!--Add species common name and scientific name to header-->

@@

```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))} 
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n"))
RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
#cat(paste("stop-time: ",Sys.time(),"\n")) 
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_ff,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
```

@@

```{r tab_ff_bestcop, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$info_ord_copname[,,1], 
             format='pandoc',
             caption = "Aphid_firstflight : best fit copula family for each selective location pair\\label{tab_ff_bestcop}",
             booktabs=TRUE
             )
```

@@

Table \ref{tab_ff_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data. Table \ref{tab_ff_AICw} represents the Akaike weights of the best-fit copula for each location pair. Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_ff_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_ff_p_KS}] of the goodness of fit with atleast
`r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` % successful bootstrap. Table \ref{tab_ff_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_ff_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC. The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_ff_LTdep_AICw} and Table \ref{tab_ff_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1. These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.

>>>>

<!--Show the statistical results in a table-->
```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,c(3,2,4)]
colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
knitr::kable(s,
             format = "pandoc",
             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
             booktabs=TRUE)
```

@@

## Results: aphid first flight data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}

@@

<!--Do the analysis-->
```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))} 
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n"))
RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
#cat(paste("stop-time: ",Sys.time(),"\n")) 
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_ff,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
```

@@

As in section \ref{Ex_count}, results are presented in $10 \times 10$ matrix format with rows
and columns corresponding to locations. Table \ref{tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the

several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data. Table \ref{tab_ff_AICw} represents the Akaike weights of the best-fit copula for each location pair. Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_ff_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_ff_p_KS}] of the goodness of fit with atleast
`r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` % successful bootstrap. Table \ref{tab_ff_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_ff_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC. The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_ff_LTdep_AICw} and Table \ref{tab_ff_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1. These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.

```{r tab_ff_bestcop, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$info_ord_copname[,,1], 
             format='pandoc',
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid. Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in Methods. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_ff_bestcop}",
             booktabs=TRUE
             )
```

@@
